---
title: "Required software"
author:
  - name: Frédéric Santos
    affiliation: CNRS, UMR 5199 PACEA
    url: https://gitlab.com/f-santos
    orcid_id: 0000-0003-1445-3871
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

The practical applications will be (partly) performed on your personal computer. Before the first day of class, make sure you have installed the following software. Also, please come with your mouse if you have one (using 3D imaging software with a trackpad is really uncomfortable).

## Office software
To handle, visualize or edit the data files, you will need:

- a spreadsheet (such as [LibreOffice Calc](https://www.libreoffice.org/download/download/), that we strongly recommend)
- a good text editor (e.g., [Notepad++](https://notepad-plus-plus.org/) for Windows, [Atom](https://atom.io/) or [SublimeText](http://www.sublimetext.com/download) for other platforms).

## MorphoJ
[MorphoJ](https://morphometrics.uk/MorphoJ_page.html) is a free software for geometric morphometrics. It is available on all operating systems.

## 3DSlicer
[3DSlicer](https://download.slicer.org/) is a free software for imaging research. It is available on all operating systems.

## Viewbox (Optional)
[Viewbox](http://www.dhal.com/download.htm) is a *non-free* software for imaging research. A limited version is available as a trial, and only for Windows users. For this reason, this download is optional (but still recommended if you can, as it will be used for the course about landmarks acquisition; see schedule for Day 3).

## Git
### Download and install Git
[Git](https://git-scm.com/downloads) will be necessary to install some R packages below.

### Configuration
Furthermore, to be able to download ("*clone*") Git repositories online, you will have to configure Git correctly.

1. **Generate a SSH key**. In a terminal (for Linux or Mac OS), or in a Git Bash terminal (for Windows), execute the following command:
   ```
   ssh-keygen -t rsa -C "votre.email@univ.fr"
   ```
   where you just have to replace `votre.email@univ.fr` by your own email. **You must use here your email from the University of Bordeaux.**
   Accept the default choice for the storage of the SSH key (proposed in the terminal by `Enter file in which to save the key`). Do *not* choose any passphrase, for it might make the subsequent steps more difficult. The process ends correctly if you get the final *randomart*.
2. **Copy the SSH key in your clipboard**. Depending on your operating system:
   - Linux users must execute the following command in a terminal:
     ```
     xclip -sel clip < ~/.ssh/id_rsa.pub
     ```
   - Mac OS users must execute the following command in a terminal:
     ```
     tr -d '\n' < ~/.ssh/id_rsa.pub | pbcopy
     ```
   - Windows users must execute the following command in a Git Bash terminal:
     ```
     cat ~/.ssh/id_rsa.pub | clip
     ```
3. **Add this SSH key to your GitUB account**. Log in to [your GitUB account](https://gitub.u-bordeaux.fr/users/sign_in) using the same username and password as for the ENT or UB Webmail. Then, visit your personal menu (topright), and click on Preferences. In the new webpage, click on SSH keys in the navigation bar on the left. In the field "Key", just paste your SSH key copied at the previous step. Choose an arbitrary title for the SSH key, and then click on the button "Add key".
4. **Check that everything is okay**. To check that your SSH key is correctly taken into account, execute the following command in a terminal (Linux or Mac OS), or in a Git Bash terminal (Windows):
   ```
   ssh -T git@gitub.u-bordeaux.fr
   ```
   If you get the message "Welcome to GitLab!", the install process was done correctly. (You might also get a warning message about "authorized administrators UB only", which is meaningless.)

## R
Please make sure that you have the *latest* version of R installed on your computer (>= 4.2.0). It can be downloaded from the [official R website](https://cran.r-project.org/).

If you currently have an older version installed, please uninstall it and download the latest version instead.

## Complementary build environment (for Mac or Windows users only)
- Mac OS: install [XCODE](https://developer.apple.com/xcode/) and [Xquartz](https://www.xquartz.org/).
- Windows: install [Rtools](https://cran.r-project.org/bin/windows/Rtools/rtools42/rtools.html) in its latest version (Rtools 4.2). 

Note that Linux users have nothing to install for this step.

## Rstudio
[Rstudio](https://www.rstudio.com/products/rstudio/download/#download) is an integrated development environment available for all operating systems.

## R packages
Run the following commands into the R console to install some useful packages:

```{r eval=FALSE, echo=TRUE, warning=FALSE}
install.packages(c("ade4", "car", "corrplot", "devtools", "epiR",
                   "factoextra", "FactoMineR", "geomorph", "ggpubr",
                   "irr", "lattice", "MASS", "Morpho", "rgl", "Rvcg",
                   "tidyverse"),
                 dep = TRUE, repos = "https://cran.wu.ac.at/")
remotes::install_git('https://gitlab.com/f-santos/anthrostat.git')
remotes::install_github("zarquon42b/mesheR")
remotes::install_github("MomX/Momocs")
```

## Help!
For any assistance, please contact me (F. Santos) directly by e-mail.
