---
title: "Required software"
author:
  - name: Frédéric Santos
    affiliation: CNRS, UMR 5199 PACEA
    url: https://gitlab.com/f-santos
    orcid_id: 0000-0003-1445-3871
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

The practical applications will be (partly) performed on your personal computer. Before the first day of class, make sure you have installed the following software.

## Usual office software
To handle, visualize or edit the data files, you will need a spreadsheet (such as [LibreOffice Calc](https://www.libreoffice.org/download/download/), that we strongly recommend), and a good text editor (e.g., [Notepad++](https://notepad-plus-plus.org/) for Windows, [Atom](https://atom.io/) or [SublimeText](http://www.sublimetext.com/download) for other platforms).

## Git
Although [Git](https://git-scm.com/downloads) will not be used by itself, it will be necessary to install some R packages below.

## R
Please make sure that you have the *latest* version of R installed on your computer. It can be downloaded from the [official R website](https://cran.r-project.org/).

## Complementary build environment (for Mac or Windows users only)
- Mac OS: install [XCODE](https://developer.apple.com/xcode/) and [Xquartz](https://www.xquartz.org/).
- Windows: install [Rtools](https://cran.r-project.org/bin/windows/Rtools/rtools42/rtools.html). 

Note that Linux users have nothing to install for this step.

## R packages
Run the following commands into the R console to install some useful packages:

```{r eval=FALSE, echo=TRUE, warning=FALSE}
install.packages(c("car", "corrplot", "devtools", "epiR",
                   "factoextra", "FactoMineR", "geomorph", "ggpubr",
                   "irr", "lattice", "MASS", "Morpho", "tidyverse"),
                 dep = TRUE, repos = "https://cran.wu.ac.at/")
devtools::install_github("MomX/Momocs")
```

## Help!
For any assistance, please contact me (F. Santos) directly by e-mail.
